\section{Handling MFA at KTH}\label{MFAsessionkeeper}% ===> this file was generated automatically by noweave --- better not edit it

Many universities, including KTH, have introduced Multi-Factor
Authentication (MFA) using Microsoft Authenticator with number matching.
This breaks the standard {\Tt{}weblogin\nwendquote} approach: the login page is now
heavily JavaScript-driven and requires a real browser to render the MFA
challenge.

The solution is a \emph{session keeper}: a script that uses a headless
Chrome browser (via Selenium) to perform the login once, handle the MFA
challenge interactively, and save the resulting cookies to a file.
Other scripts can then load these cookies and make requests without
going through the login flow again.

The module looks like this.
\nwfilename{mfa.nw}\nwbegincode{1}\sublabel{NWXRURf-1Fx3qe-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-1Fx3qe-1}}}\moddef{\code{}mfa.py\edoc{}~{\nwtagstyle{}\subpageref{NWXRURf-1Fx3qe-1}}}\endmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
import getpass
import http.cookiejar
import os
import pickle
import re
import sys

from selenium import webdriver
from selenium.webdriver.chrome.options import Options
from selenium.webdriver.common.by import By
from selenium.webdriver.support import expected_conditions as EC
from selenium.webdriver.support.ui import WebDriverWait

import requests.cookies

\LA{}set up headless browser~{\nwtagstyle{}\subpageref{NWXRURf-ipLdC-1}}\RA{}
\LA{}perform login~{\nwtagstyle{}\subpageref{NWXRURf-46xhjq-1}}\RA{}
\LA{}find MFA number~{\nwtagstyle{}\subpageref{NWXRURf-33O1O9-1}}\RA{}
\LA{}collect cookies~{\nwtagstyle{}\subpageref{NWXRURf-4ZzA54-1}}\RA{}
\LA{}save cookies~{\nwtagstyle{}\subpageref{NWXRURf-1N6piN-1}}\RA{}
\LA{}command-line interface~{\nwtagstyle{}\subpageref{NWXRURf-3jIocA-1}}\RA{}
\nwnotused{[[mfa.py]]}\nwendcode{}\nwbegindocs{2}\nwdocspar

\subsection{Setting up the headless browser}

We use Chrome in headless mode.
The {\Tt{}--no-sandbox\nwendquote} and {\Tt{}--disable-dev-shm-usage\nwendquote} flags are needed
when running on a server without a display.
\nwenddocs{}\nwbegincode{3}\sublabel{NWXRURf-ipLdC-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-ipLdC-1}}}\moddef{set up headless browser~{\nwtagstyle{}\subpageref{NWXRURf-ipLdC-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-1Fx3qe-1}}\nwenddeflinemarkup
def setup_browser():
  """Start a headless Chrome browser and return the driver."""
  options = Options()
  options.add_argument("--headless=new")
  options.add_argument("--no-sandbox")
  options.add_argument("--disable-dev-shm-usage")
  options.add_argument("--disable-gpu")
  options.add_argument("--window-size=1280,900")
  return webdriver.Chrome(options=options)
\nwused{\\{NWXRURf-1Fx3qe-1}}\nwendcode{}\nwbegindocs{4}\nwdocspar

\subsection{Performing the login}

The login flow is as follows.
\begin{enumerate}
\item Navigate to a URL that triggers a redirect to the KTH login page.
\item Fill in credentials using JavaScript injection.
  We cannot use {\Tt{}send{\_}keys\nwendquote} because it fails with
  \enquote{element not interactable} in headless Chrome on this page.
\item Wait for either the MFA number to appear or for the login to
  complete directly (some services skip MFA on trusted networks).
\item If MFA is required, print the number and wait for the user to
  approve on their phone.
\end{enumerate}
\nwenddocs{}\nwbegincode{5}\sublabel{NWXRURf-46xhjq-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-46xhjq-1}}}\moddef{perform login~{\nwtagstyle{}\subpageref{NWXRURf-46xhjq-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-1Fx3qe-1}}\nwprevnextdefs{\relax}{NWXRURf-46xhjq-2}\nwenddeflinemarkup
def do_login(driver, username, password, trigger_url):
  """
  Navigate to trigger_url, fill credentials, handle MFA if needed.
  Raises RuntimeError on timeout.
  """
  driver.get(trigger_url)

  \LA{}wait for login page~{\nwtagstyle{}\subpageref{NWXRURf-lIcJE-1}}\RA{}
  \LA{}fill credentials~{\nwtagstyle{}\subpageref{NWXRURf-45NdAu-1}}\RA{}
  \LA{}wait for MFA or completion~{\nwtagstyle{}\subpageref{NWXRURf-rpbYU-1}}\RA{}
\nwalsodefined{\\{NWXRURf-46xhjq-2}}\nwused{\\{NWXRURf-1Fx3qe-1}}\nwendcode{}\nwbegindocs{6}\nwdocspar

We wait until we land on {\Tt{}login.ug.kth.se\nwendquote}, which is KTH's
authentication server.
\nwenddocs{}\nwbegincode{7}\sublabel{NWXRURf-lIcJE-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-lIcJE-1}}}\moddef{wait for login page~{\nwtagstyle{}\subpageref{NWXRURf-lIcJE-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-46xhjq-1}}\nwenddeflinemarkup
WebDriverWait(driver, 20).until(
  lambda d: "login.ug.kth.se" in d.current_url)
\nwused{\\{NWXRURf-46xhjq-1}}\nwendcode{}\nwbegindocs{8}\nwdocspar

We fill in credentials via JavaScript because {\Tt{}send{\_}keys\nwendquote} does not
work reliably in headless Chrome on this particular page.
We append {\Tt{}@ug.kth.se\nwendquote} if needed, since the login form expects the
full UG username.
\nwenddocs{}\nwbegincode{9}\sublabel{NWXRURf-45NdAu-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-45NdAu-1}}}\moddef{fill credentials~{\nwtagstyle{}\subpageref{NWXRURf-45NdAu-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-46xhjq-1}}\nwenddeflinemarkup
uname = username if "@ug.kth.se" in username \\
                 else username + "@ug.kth.se"
driver.execute_script(
  "document.getElementById('userNameInput').value = arguments[0];",
  uname)
driver.execute_script(
  "document.getElementById('passwordInput').value = arguments[0];",
  password)
driver.execute_script(
  "document.getElementById('submitButton').click();")
\nwused{\\{NWXRURf-46xhjq-1}}\nwendcode{}\nwbegindocs{10}\nwdocspar

After submitting, we may land on an MFA challenge or proceed directly
to the target page.
We wait for either condition.
\nwenddocs{}\nwbegincode{11}\sublabel{NWXRURf-rpbYU-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-rpbYU-1}}}\moddef{wait for MFA or completion~{\nwtagstyle{}\subpageref{NWXRURf-rpbYU-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-46xhjq-1}}\nwenddeflinemarkup
def mfa_or_done(d):
  number = find_mfa_number(d)
  if number:
    return ("mfa", number)
  if not _on_login_page(d.current_url):
    return ("done", None)
  return None

result = WebDriverWait(driver, 30).until(mfa_or_done)

if result[0] == "mfa":
  \LA{}handle MFA challenge~{\nwtagstyle{}\subpageref{NWXRURf-4J8dR3-1}}\RA{}
\nwused{\\{NWXRURf-46xhjq-1}}\nwendcode{}\nwbegindocs{12}\nwdocspar

When MFA is required, we print the number and wait up to two minutes
for the user to approve on their phone.
\nwenddocs{}\nwbegincode{13}\sublabel{NWXRURf-4J8dR3-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-4J8dR3-1}}}\moddef{handle MFA challenge~{\nwtagstyle{}\subpageref{NWXRURf-4J8dR3-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-rpbYU-1}}\nwenddeflinemarkup
number = result[1]
print("\\n" + "=" * 50)
print("  Enter  \{\}  in Microsoft Authenticator".format(number))
print("=" * 50 + "\\n")
WebDriverWait(driver, 120).until(
  lambda d: not _on_login_page(d.current_url))
\nwused{\\{NWXRURf-rpbYU-1}}\nwendcode{}\nwbegindocs{14}\nwdocspar

We use a small helper to check if we are still on the login page.
\nwenddocs{}\nwbegincode{15}\sublabel{NWXRURf-46xhjq-2}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-46xhjq-2}}}\moddef{perform login~{\nwtagstyle{}\subpageref{NWXRURf-46xhjq-1}}}\plusendmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-1Fx3qe-1}}\nwprevnextdefs{NWXRURf-46xhjq-1}{\relax}\nwenddeflinemarkup
def _on_login_page(url):
  return "login.ug.kth.se" in url or "microsoftonline.com" in url
\nwused{\\{NWXRURf-1Fx3qe-1}}\nwendcode{}\nwbegindocs{16}\nwdocspar

\subsection{Finding the MFA number}

Microsoft Authenticator number matching displays a two-digit number
that the user must enter on their phone.
The number appears in a specific element, but the element ID has varied
across Microsoft updates.
We therefore try several known IDs and fall back to scanning the page
body for any standalone two-digit number.
\nwenddocs{}\nwbegincode{17}\sublabel{NWXRURf-33O1O9-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-33O1O9-1}}}\moddef{find MFA number~{\nwtagstyle{}\subpageref{NWXRURf-33O1O9-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-1Fx3qe-1}}\nwenddeflinemarkup
_MFA_ELEMENT_IDS = [
  "idRichContext_DisplaySign",
  "displaySign",
  "idDiv_DisplaySign",
]

def find_mfa_number(driver):
  """
  Return the MFA number shown on the current page, or None if not present.
  """
  for eid in _MFA_ELEMENT_IDS:
    elems = driver.find_elements(By.ID, eid)
    if elems:
      txt = elems[0].text.strip()
      if txt:
        return txt
  \LA{}fall back to body scan~{\nwtagstyle{}\subpageref{NWXRURf-2FUw1V-1}}\RA{}
\nwused{\\{NWXRURf-1Fx3qe-1}}\nwendcode{}\nwbegindocs{18}\nwdocspar

If no known element is found, we scan the visible body text.
This is a heuristic, but in practice the MFA number is the only
prominent two-digit number on the page.
\nwenddocs{}\nwbegincode{19}\sublabel{NWXRURf-2FUw1V-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-2FUw1V-1}}}\moddef{fall back to body scan~{\nwtagstyle{}\subpageref{NWXRURf-2FUw1V-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-33O1O9-1}}\nwenddeflinemarkup
body = driver.find_element(By.TAG_NAME, "body").text
matches = re.findall(r'\\b(\\d\{2\})\\b', body)
return matches[0] if matches else None
\nwused{\\{NWXRURf-33O1O9-1}}\nwendcode{}\nwbegindocs{20}\nwdocspar

\subsection{Collecting cookies}

After a successful login, we collect all cookies from the browser and
convert them into a {\Tt{}requests.cookies.RequestsCookieJar\nwendquote}.
We use {\Tt{}http.cookiejar.Cookie\nwendquote} objects directly rather than
{\Tt{}jar.set()\nwendquote}, because {\Tt{}jar.set()\nwendquote} does not correctly handle domain
cookies with a leading dot (e.g.\ {\Tt{}.kth.se\nwendquote}).
\nwenddocs{}\nwbegincode{21}\sublabel{NWXRURf-4ZzA54-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-4ZzA54-1}}}\moddef{collect cookies~{\nwtagstyle{}\subpageref{NWXRURf-4ZzA54-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-1Fx3qe-1}}\nwenddeflinemarkup
def collect_cookies(driver, extra_urls=None):
  """
  Collect cookies from the browser and return a RequestsCookieJar.
  If extra_urls is given, also visit those URLs to pick up their cookies
  (useful when SSO spans multiple domains).
  """
  jar = requests.cookies.RequestsCookieJar()
  _add_driver_cookies(jar, driver)

  for url in (extra_urls or []):
    driver.get(url)
    WebDriverWait(driver, 30).until(
      lambda d: not _on_login_page(d.current_url))
    _add_driver_cookies(jar, driver)

  return jar

def _add_driver_cookies(jar, driver):
  for c in driver.get_cookies():
    domain = c.get("domain", "")
    cookie = http.cookiejar.Cookie(
      version=0,
      name=c["name"],
      value=c["value"],
      port=None, port_specified=False,
      domain=domain,
      domain_specified=bool(domain),
      domain_initial_dot=domain.startswith("."),
      path=c.get("path", "/"),
      path_specified=bool(c.get("path")),
      secure=c.get("secure", False),
      expires=c.get("expiry"),
      discard=False,
      comment=None, comment_url=None,
      rest=\{"HttpOnly": ""\} if c.get("httpOnly") else \{\},
    )
    jar.set_cookie(cookie)
\nwused{\\{NWXRURf-1Fx3qe-1}}\nwendcode{}\nwbegindocs{22}\nwdocspar

\subsection{Saving cookies}

We save the cookie jar to a file using {\Tt{}pickle\nwendquote}, so that other
scripts can load it with {\Tt{}pickle.load\nwendquote} and pass the jar directly to
a {\Tt{}requests.Session\nwendquote}.
\nwenddocs{}\nwbegincode{23}\sublabel{NWXRURf-1N6piN-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-1N6piN-1}}}\moddef{save cookies~{\nwtagstyle{}\subpageref{NWXRURf-1N6piN-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-1Fx3qe-1}}\nwenddeflinemarkup
def save_cookies(jar, cookie_file):
  """Save a RequestsCookieJar to cookie_file using pickle."""
  os.makedirs(os.path.dirname(os.path.abspath(cookie_file)), exist_ok=True)
  with open(cookie_file, "wb") as f:
    pickle.dump(jar, f)
\nwused{\\{NWXRURf-1Fx3qe-1}}\nwendcode{}\nwbegindocs{24}\nwdocspar

\subsection{Command-line interface}

When run as a script ({\Tt{}python\ -m\ weblogin.mfa\nwendquote}), the module asks for
credentials and a target URL, performs the login, and saves the cookies.
\nwenddocs{}\nwbegincode{25}\sublabel{NWXRURf-3jIocA-1}\nwmargintag{{\nwtagstyle{}\subpageref{NWXRURf-3jIocA-1}}}\moddef{command-line interface~{\nwtagstyle{}\subpageref{NWXRURf-3jIocA-1}}}\endmoddef\nwstartdeflinemarkup\nwusesondefline{\\{NWXRURf-1Fx3qe-1}}\nwenddeflinemarkup
def main():
  """
  Perform a KTH MFA login and save cookies to a file.

  Usage: python -m weblogin.mfa <trigger_url> <cookie_file>
  """
  if len(sys.argv) < 3:
    print("Usage: python -m weblogin.mfa <trigger_url> <cookie_file>",
          file=sys.stderr)
    sys.exit(1)

  trigger_url = sys.argv[1]
  cookie_file = sys.argv[2]

  username = input("KTH username: ")
  password = getpass.getpass("KTH password: ")

  driver = setup_browser()
  try:
    do_login(driver, username, password, trigger_url)
    jar = collect_cookies(driver)
    save_cookies(jar, cookie_file)
    print("Cookies saved to \{\}".format(cookie_file))
  except Exception as e:
    print("ERROR: \{\}".format(e), file=sys.stderr)
    sys.exit(1)
  finally:
    driver.quit()

if __name__ == "__main__":
  main()
\nwused{\\{NWXRURf-1Fx3qe-1}}\nwendcode{}

\nwixlogsorted{c}{{\code{}mfa.py\edoc{}}{NWXRURf-1Fx3qe-1}{\nwixd{NWXRURf-1Fx3qe-1}}}%
\nwixlogsorted{c}{{collect cookies}{NWXRURf-4ZzA54-1}{\nwixu{NWXRURf-1Fx3qe-1}\nwixd{NWXRURf-4ZzA54-1}}}%
\nwixlogsorted{c}{{command-line interface}{NWXRURf-3jIocA-1}{\nwixu{NWXRURf-1Fx3qe-1}\nwixd{NWXRURf-3jIocA-1}}}%
\nwixlogsorted{c}{{fall back to body scan}{NWXRURf-2FUw1V-1}{\nwixu{NWXRURf-33O1O9-1}\nwixd{NWXRURf-2FUw1V-1}}}%
\nwixlogsorted{c}{{fill credentials}{NWXRURf-45NdAu-1}{\nwixu{NWXRURf-46xhjq-1}\nwixd{NWXRURf-45NdAu-1}}}%
\nwixlogsorted{c}{{find MFA number}{NWXRURf-33O1O9-1}{\nwixu{NWXRURf-1Fx3qe-1}\nwixd{NWXRURf-33O1O9-1}}}%
\nwixlogsorted{c}{{handle MFA challenge}{NWXRURf-4J8dR3-1}{\nwixu{NWXRURf-rpbYU-1}\nwixd{NWXRURf-4J8dR3-1}}}%
\nwixlogsorted{c}{{perform login}{NWXRURf-46xhjq-1}{\nwixu{NWXRURf-1Fx3qe-1}\nwixd{NWXRURf-46xhjq-1}\nwixd{NWXRURf-46xhjq-2}}}%
\nwixlogsorted{c}{{save cookies}{NWXRURf-1N6piN-1}{\nwixu{NWXRURf-1Fx3qe-1}\nwixd{NWXRURf-1N6piN-1}}}%
\nwixlogsorted{c}{{set up headless browser}{NWXRURf-ipLdC-1}{\nwixu{NWXRURf-1Fx3qe-1}\nwixd{NWXRURf-ipLdC-1}}}%
\nwixlogsorted{c}{{wait for login page}{NWXRURf-lIcJE-1}{\nwixu{NWXRURf-46xhjq-1}\nwixd{NWXRURf-lIcJE-1}}}%
\nwixlogsorted{c}{{wait for MFA or completion}{NWXRURf-rpbYU-1}{\nwixu{NWXRURf-46xhjq-1}\nwixd{NWXRURf-rpbYU-1}}}%
\nwbegindocs{26}\nwdocspar
\nwenddocs{}
